Proposal to improve BBC Search through the use of Linked Data

### 2014-05-19 chapter1.tex
Did some reviewing and trrying to fit diagrams in.

\begin{enumerate}
  \item Further decoupling of BBC applications, which aids scalability,
    maintainability and resilience in many software architectures.\cite{}
  \item Improvements to the search results page that help users determine
    if the results given are relevant via so-called
    \emph{Rich Snippets}\cite{goel2009introducing}
    that can automatically display some semantic information about a given
    result without the need for manual curation.
  \item User experience and design improvements to the search results
    where semantic properties about the results can drive visual decoration
    and calls to action, which can allow for a more flexible and
    adaptable design without the need to enumerate all ``kinds'' with
    different templates.
  \item Removal of the manual curation of the ``Editor's Choice'' results
    against a given query and replacing it with curated linked data that
    can provide recommended content against queries in a more general
    manner.
\end{enumerate}

In the rest of this chapter, I will give an overview of what role BBC
Search plays and how we can usefully view the system as four
distinct subsystems. In later chapters we can then look to improve these
four separate systems one at a time and then evaluate how these improvements
might complement each other.

@@

In the remaining sections of this chapter, I outline the considerations
around these two sides of the search application in addition to the
two complementary systems known as \emph{Editor's Choice} and
\emph{Search Suggest}.

% TODO

@@

Note that a system could trivially maximise recall by always returning all
results, but such a system would perform poorly in terms of precision. A
measure known as \emph{F-measure} was derived by van Rijsbergen
\cite{rijsbergen1979information} which
balances precision recall based on preferring recall over precision $\beta$
times:

\begin{displaymath}
  F_\beta = (1 + \beta^2) \cdot \frac{\mathrm{precision} \cdot \mathrm{recall} }{ \beta^2 \cdot \mathrm{precision} + \mathrm{recall}}
\end{displaymath}

There are other measures such as average precision and average mean precision
that take into account ordering of results and the average over multiple
queries respectively. A successful algorithm or third party search engine
application can be evaluated quantitatively by these measures and more. A more
qualitative summary from the perspective of the end users could be:

\begin{quote}
  A successful search application will return only the content the user wants
  (precision) --
  as high up the results list as possible -- and as much of desired content
  as possible (recall).
\end{quote}

@@

\section{Summary}

In chapter~\ref{stateoftheart} I will briefly outline some of the state of the
art in the Linked Data domain and start to highlight areas that show
the most promise for improving the four areas of the search application
described in this chapter.

>>>>

\begin{enumerate}
  \item Further decoupling of BBC applications, which aids scalability,
    maintainability and resilience in many software architectures.
  \item Improvements to the search results page that help users determine
    if the results given are relevant. The example of this is by using
    \emph{Rich Snippets}\cite{goel2009introducing}
    that can automatically display some semantic information about a given
    result without the need for manual curation.
  \item User experience and design improvements to the search results
    where semantic properties about the results can drive visual decoration
    and calls to action.
  \item Removal of the exclusively manual curation of the ``Editor's Choice'' results
    against a given query and replacing it with curated linked data that
    can provide recommended content against queries in a more general
    manner.
\end{enumerate}

In the rest of this chapter, I will give an overview of what role BBC
Search plays and the distinct subsystems that it comprises.
In later chapters we can then look to improve these systems
how these improvements might complement each other.

@@

@@

Note that a system could trivially maximise recall by always returning all
results, but such a system would perform poorly in terms of precision.
This makes a search application a delicate balance of trying to show
users as much of what they want as possible without cluttering the results
with irrelevant items.

The task is complicated somewhat further from a user experience perspective
if an application wishes to attempt to create \emph{serendipity}. The
concept of serendipity -- showing the user something they were not looking
for, but turns into a positive onward journey -- is recognised
as part of the search user experience and is something human-computer
interaction designers try to aspire to create.\cite{russell2012designing}

@@

### 2014-05-19 chapter2.tex
Did some reviewing and trrying to fit diagrams in.

\begin{enumerate}
  \item Use URIs to denote things.
  \item Prefer HTTP URIs (i.e. URLs) so that these things can be looked
    up (i.e. \emph{dereferenced}.
  \item When these URIs are dereferenced, provide useful information about
    those things using standards.
  \item Include links to other things (using their URIs) when such
    information is published on the web.
\end{enumerate}

It can be seen as a subset of Semantic Web techniques and it builds upon
the \emph{Resource Description Framework} (RDF)\cite{lassila1999resource},
which is a general-purpose method for describing and modelling information.

In this chapter, I will outline how web authors are encouraged to publish
linked data within their content and how that is already used to enhance
search results in global search engines. I will then introduce how linked
data may be used within an enterprise setting to reduce traditional
integration efforts and then touch on some other aspects of the linked
data domain that could be of interest to the search application.

@@

\begin{itemize}
  \item RDFa\cite{adida2012rdfa} and RDFa Lite\cite{lite2004rdfa}.
  \item DC-HTML defines how to embed the Dublin Core vocabulary within HTML
    meta tags.
  \item Microdata is an HTML standard that allows annotating existing markup
    via additional attributes so that it serves as metadata in addition to
    its purpose as text for humans.
  \item Microformats are another mechanism that embeds semantic metadata
    within existing HTML tags and attributes.
\end{itemize}

@@

Google have seemingly partially adopted an approach following this principle
-- at least for web authors publishing information about music
events.\cite{googlejsonld} Google are piloting embedding RDF metadata in
\emph{JSON-LD} format within HTML \texttt{<script>} tags to add semantic
information alongside textual data, but without annotating this metadata
around the textual content itself as we would with RDFa, for example.

All of the above approaches can be found ``in the wild'' in pages across
the Web with differing rates of adoption. General-purpose web search
engines can make use of this additional metadata to improve the quality
of the search results in a number of ways. In section~\ref{enhancing-results},
I discuss some of these potential improvements.

@@

@@

A higher CTR for semantically-enhanced search results creates some pressure
and incentive on web publishers to include semantic metadata in their web
content if they aspire to attacting visitors via global search engines such
as Google, Bing and Yahoo!.

@@

In recent years, a growing trend has emerged to use
Linked Data in the enterprise to remove the need
for repeated integration projects or data warehousing.
indexes over bespoke integration efforts. This follows from
a major motivation of Linked Data being that it aids in breaking
down information silos.\cite{bizer2009linked} In the
\emph{Linked Enterprise Data} approach, all information
production remains \emph{distributed} but is
\emph{connectable}.\cite{allemang2010semantic}

An organisation that seeks to ensure that all information is
published as Linked Data can learn from examples such as Wikipedia
-- or indeed the web in general -- where individual experts are
encouraged to share knowledge rather than keep it in silos. This
distributed approach recognises that diversity and flux are the
steady state of an agile enterprise. Two sources that happen
to use the same vocabulary are instantly interoperable. Two
sources that use a different vocabulary could well be integrated
simply by creating appropriate mappings between the vocabularies.
It should not be necessary to re-engineer a whole dataset nor
create entirely new systems just for integration.

@@

\begin{comment}
If we view the indexing processes
of a search application as a variant of data warehousing,
we can start to argue for enterprise-wide publishing of
linked data as a more cost effective means of building search

@@

These efforts are not as structured and rich as the more notable uses
of semantic web technologies on the programmes microsites, Nature and BBC
Music \cite{raimond2010use}, but indicate that some level of metadata
can be retrieved
as linked data. In cases where the metadata are rich enough, it might be
possible to avoid direct integration with databases, hidden back end
web services and other structured ``feeds'' typically set up for enterprise
integration.
\end{comment}

>>>>

\begin{enumerate}
\item Use URIs to denote things, whether pages/documents on the web
  or real physical and conceptual things in the world.
\item Prefer HTTP URIs (i.e. URLs) so that they can be looked
  up (i.e. \emph{dereferenced} to get more information about the thing denoted.
\item When these URIs are dereferenced, provide useful information about
  those things using web standards, e.g. RDF, HTML, XML, Turtle.
\item Include links to other things (using their URIs) when such
  information is published on the web.
\end{enumerate}

This overlaps somewhat Semantic Web techniques and it builds upon
the \emph{Resource Description Framework} (RDF)\cite{lassila1999resource},
which is a general-purpose method for describing and modelling information.

In this chapter, I will outline how web authors are encouraged to publish
linked data within their content and how that is already used to enhance
search results in global search engines. I will then introduce how linked
data may be used within an enterprise setting to reduce traditional
integration efforts Finally, I will then touch on some other aspects of the linked
data domain that could be of interest to the search application.

@@

\begin{itemize}
  \item RDFa\cite{adida2012rdfa} and RDFa Lite\cite{lite2004rdfa}.
  \item DC-HTML defines how to embed the Dublin Core vocabulary within HTML
    meta tags.\cite{johnston2008expressing}
  \item Microdata\cite{htmlstandard} is an HTML standard that allows
    annotating existing markup
    via additional attributes so that it serves as metadata in addition to
    its purpose as text for humans.
  \item Microformats are another mechanism that embeds semantic metadata
    within existing HTML tags and attributes.
\end{itemize}

@@

Google have seemingly partially adopted an approach following this principle
-- at least for web authors publishing information about music
events.\cite{googlejsonld} Google are piloting embedding RDF metadata in
\emph{JSON-LD} format within HTML \texttt{<script>} tags to add semantic
information alongside textual data, but without needing to annotate this metadata
around the textual content itself as we would with RDFa, for example.

All of the above approaches can be found ``in the wild'' in pages across
the Web with differing rates of adoption.\cite{haas2011enhanced}
General-purpose web search
engines can make use of this additional metadata to improve the quality
of the search results in a number of ways. In section~\ref{enhancing-results},
I discuss some of these potential improvements.

@@

This is out of scope for this paper, but
is recommended as a further study after this proposal has been implemented.

@@

A higher CTR for semantically-enhanced search results creates some pressure
and incentive on web publishers to include semantic metadata in their web
content if they aspire to attracting visitors via global search engines such
as Google, Bing and Yahoo!.

@@

In recent years, a growing trend has emerged to use
Linked Data in the enterprise to remove the need
for repeated integration projects or data warehousing.
indexes over bespoke integration efforts. This follows from
a major motivation of Linked Data being that it aids in breaking
down information silos.\cite{bizer2009linked} The key theme
in the \emph{Linked Enterprise Data} approach is that all information
production remains \emph{distributed} but becomes
\emph{connectable}.\cite{allemang2010semantic}

A distributed
approach to such publishing across an enterprise keeps data sources
interoperable, but allows them to develop and grow independantly. This
recognises that such flux and diversity is the steady state of
an agile enterprise and shows that attempts to impose a controlled
data store and vocabulary across the enterprise are doomed to fail.
Integration of two previously-disconnected data sources is simply
a matter of creating appropriate vocabulary mappings. It is no
longer necessarily to build bespoke middleware or rebuild one or
both data stores.

@@

\section{Linked Data and Search}

If we view the indexing processes
of a search application as a variant of data warehousing,
we can start to argue for enterprise-wide publishing of
linked data as a more cost effective means of building search

@@

These efforts are not as structured and rich as the more notable uses
of semantic web technologies on the programmes microsites, Nature and BBC
Music \cite{raimond2010use}, but indicate that some level of metadata
can be retrieved
as linked data. In cases where the metadata are rich enough, it might be
possible to avoid direct integration with databases, hidden back end
web services and other structured ``feeds'' typically set up for enterprise
integration.

### 2014-05-19 chapter4.1.tex
Almost finished chapter 4.
Did some reviewing and trrying to fit diagrams in.

\section{Linked Data Indexing}

@@

>>>>

\section{Linked Data Indexing}
\label{sec:ld-indexing}

@@

Whilst the enrichment stage means we have not fully eliminated the
need for bespoke integration work, the important factor to consider is that
such integration is no longer \emph{required}. If basic properties
are sourced upstream from the public web page, then the lack
of ability to integration -- or if that integration breaks, as brittle,
coupled integration applications are prone to do -- then the information
can still be passed through to retain a \emph{degraded} search
experience. This is preferable to the complete loss of an upstream feed
due to incompatible changes being made unannounced.

### 2014-05-19 chapter5.3.tex
Did some reviewing and trrying to fit diagrams in.

\begin{figure}
  \begin{center}
    \begin{dot2tex}[dot,pgf,scale=0.41]
      \input{basic-ontology.dot}
    \end{dot2tex}
  \end{center}
  \caption{Basic ontology modelling the traditional ``title, link, description'' search result}
  \label{fig:initial-ontology}
\end{figure}

A common model across differing BBC content items has been
attempted\cite{fenning2014applicability} at a high level, but
for first iterations of this feature, a simplistic model
such as that shown in figure~\ref{fig:initial-onotology}
would suffice.

@@

The ``Ingest'' subteam needs to ensure the Ingest API can
accept content meeting the basic vocabulary outline in
figure~\ref{fig:initial-onotologu}, validating ones
that do not contain the basic fields and evolve the API. If
the results page developers choose to alter the domain
model based on design iterations and user feedback, then
these changes can propogate back as requirements for the
content ingest.

@@

\begin{lstlisting}[language=Python]
from elasticsearch import Elasticsearch
import json

def ingest_jsonld(jsonld_string):
  body = {'jsonld': json.loads(jsonld_string)}
  ElasticSearch().index(
    index='search', body=body, doc_type='item')
\label{lst:ingest}
\end{lstlisting}

>>>>

\begin{figure}
  \begin{center}
    \begin{dot2tex}[dot,pgf,scale=0.42,options=-traw]
      \input{basic-ontology.dot}
    \end{dot2tex}
  \end{center}
  \caption{Basic ontology modelling the traditional ``title, link, description'' search result\label{fig:initial-ontology}}
\end{figure}

A common model across differing BBC content items has been
attempted\cite{fenning2014applicability} at a high level, but
for first iterations of this feature, a simplistic model
such as that shown in figure~\ref{fig:initial-ontology}
would suffice.

@@

The ``Ingest'' subteam needs to ensure the Ingest API can
accept content meeting the basic vocabulary outline in
figure~\ref{fig:initial-ontology}, validating ones
that do not contain the basic fields and evolve the API. If
the results page developers choose to alter the domain
model based on design iterations and user feedback, then
these changes can propogate back as requirements for the
content ingest.

@@

\begin{lstlisting}[language=Python,label=lst:ingest]
from elasticsearch import Elasticsearch
import json

def ingest_jsonld(jsonld_string):
  body = {'jsonld': json.loads(jsonld_string)}
  ElasticSearch().index(
    index='search', body=body, doc_type='item')
\end{lstlisting}
